---
title: "Homework 2"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Referring to the data in Table below, answer the following questions.
(a) Use Student’s t-test to examine whether the difference in body weight between genotypes 1 and 0 for the marker is statistically significant.
```{r}
y = c(28, 32, 28, 29, 31, 20, 18, 21, 20, 21) # data input # space # enter
x = matrix(c(rep(1,10),rep(1,5),rep(0,5)),10,2) # 10 rows , 2 columns
trt = gl(2,5,10) # gl = Generate Factor Levels,  2 levels, 5 replications

# t-test
## test VARs are equal or not 
var.result <- var.test(y~trt) 
var.result
print(paste("p-value =",var.result$p.value))
print("No significant difference -> VARs are equal.")

cat("Line \nLine 2")


## equal variance t.test
t.test.result <- t.test(y~trt, var.equal=TRUE)
t.test.result
print(paste("p-value =",t.test.result$p.value))
print("The difference in body weight between genotypes 1 and 0 for the marker is statistically significant.")

```

(b) Perform the single marker analysis to estimate the genetic effect, a, for the marker. Provide the details (screen shot) or codes about your calculation.
```{r}
# linear algebra to estimate the genetic coefficient
b = solve( t(x) %*% x) %*% (t(x) %*% y) # inverse = FUN solve (解二元一次方程式)
b
print(paste("Genetic effect (a) =",b[1,1]))
sigma2 = t(y-x%*%b) %*% (y-x%*%b)/10
sigma2
```

(c) Perform the F-test to examine whether the marker exerts a significant effect on body weight in the backcross population of mice.
```{r}
# F-test
out = lm(y~trt) # linear model
result = anova(out) # 列出變方分析表
result
print(paste("Pr(>F) =",result$`Pr(>F)`[1]))
print("The marker exerts a significant effect.")
```

#### 2. Apply the QTL analysis on the “listeria” sample data set in r/qtl.

(a) Apply the single marker analysis (Sax, 1923) to all markers and summarize the markers which have LOD score > 3.
```{r}
# install.packages("qtl") # don't need to re-download it i
library(qtl)
?listeria # check the details of the dataset
data(listeria) # sample dataset: Susceptibility to Listeria monocytogenes infection in the mouse, 120 F2 females, 133 markers(2 on the X chromosome), phenotype = survival time (in hours) following infection

# data summary
plot(listeria)
listeria$pheno #phenotype long table
geno.table(listeria) #genotype long table

plotPXG(listeria,"D10M44") # plot one marker as an example
plotPXG(listeria,"D6M223") # plot one marker as an example
plotRF(listeria) #estimate and plot the pairwise recombinant fraction, r

# single marker analysis
## "mr" stands for "marker regression" 
## scanone = 一次只scan一個
out.mr = scanone(listeria, method="mr") 

# assume only one marker on each chromosome，且只在每個chr上選最顯著的marker
summary(out.mr)  # all markers
summary(out.mr, threshold=3) # markers which have LOD score > 3

plot(out.mr, col=1)
legend("topright",c("MR"), col=1, lty=1)

# 人為選出大致上看起來最高及最平緩的區間畫圖
plot(out.mr, chr=c(5,14)) 
legend("topright",c("MR"), col=1, lty=1)
plot(out.mr, chr=c(5,14), type="h")
legend("topright",c("MR"), col=1, lty=1)
```

#### In (b) and (c), we need:
```{r}
# interval mapping: IM (Lander; EM) or RIM (Harley-Knott; H-K)
# step: distance (in cM) between positions at which the genotype probabilities are calculated
listeria = calc.genoprob(listeria, step=1)  # 1 centimorgan , chromosome interval unit, grid search
```

(b) Perform the interval mapping (Lander and Botstein, 1989), plot the results, and summarize the significant marker associated with the trait having permutation p-values < 0.05.
```{r}
# IM (Lander; EM)
listeria = calc.genoprob(listeria, step=1)  # 1 centimorgan , chromosome interval unit, grid search
out.em = scanone(listeria, method="em")
head(out.em) # show step = 1 (start from the 1st marker) 
out.em[1:30,] #我們設定的檢查區間為1 centimorgan，若檢查區間無marker，則每次檢查區間皆為1，若檢查區間出現marker則會出現區間不為1的情況，如0-1之間的D1M3、24-25之間的D1M75。
plot(out.em, col=2)
legend("topright",c("EM"), col=2, lty=1)
plot(out.mr,out.em, chr=c(5,14),col=1:2)
legend("topright",c("MR","EM"), col=1:2, lty=1)


# permutation test for IM
operm = scanone(listeria, method="em", n.perm=1000, verbose=FALSE) # verbose=FALSE: suppress output # n.perm = permutation 
operm.f = data.frame(operm)
summary.em = summary(out.em, perms=operm, alpha=0.05, pvalues=TRUE)
summary.em 
# 由於沒有任何 marker 在 permutation 的結果中 p-values < 0.05，所以我重新設定一次，將 alpha 設為 0.1，找出以下兩個 marker。
summary.em.1 = summary(out.em, perms=operm, alpha=0.1, pvalues=TRUE)
summary.em.1 

# Plot those significant markers associated with the trait having permutation p-values < 0.1 (沒有任何 marker 在 permutation 的結果中 p-values < 0.05) 
hist(operm.f$lod,ylim=c(0,700))
text(x=6, y=500,"permutation" , pos=2)
abline(v=summary.em.1$lod,col=c("green","darkgreen"))
text(x= summary.em.1$lod[1]+4, y=650, rownames(summary.em.1)[1], pos=2,col="green")
text(x= summary.em.1$lod[2]+1, y=650, rownames(summary.em.1)[2], pos=2,col="darkgreen")

```

(c) Perform the regression interval mapping (Harley and Knott, 1992), plot the results, and summarize the significant marker associated with the trait having permutation p-values < 0.05.
```{r}
# RIM (Harley-Knott; H-K)
listeria = calc.genoprob(listeria, step=1)  # 1 centimorgan , chromosome interval unit, grid search
out.hk = scanone(listeria, method="hk") # method預設值＝ＥＭ
plot(out.hk, col=4)
legend("topright",c("HK"), col=4, lty=1)
plot(out.mr,out.em,out.hk, chr=c(5,14),col=c(1,2,4))
legend("topright",c("MR","EM","HK"), col=c(1,2,4), lty=1)

# permutation test for HK
operhk = scanone(listeria, method="hk", n.perm=1000, verbose=FALSE) # verbose=FALSE: suppress output
operhk.f = data.frame(operhk)
summary.hk = summary(out.hk, perms=operhk, alpha=0.05, pvalues=TRUE) 
summary.hk

# Plot those significant markers associated with the trait having permutation p-values < 0.05 
hist(operhk.f$lod,xlim=c(1,7))
text(x=3, y=300,"permutation" , pos=2)
abline(v=summary.hk$lod,col=c("green","darkgreen"))
text(x= summary.hk$lod[1]+0.7, y=300, rownames(summary.hk)[1], pos=2, col="green")
text(x= summary.hk$lod[2]+0.3, y=300, rownames(summary.hk)[2], pos=2, col="darkgreen")

```


(d) Compare the result of the above three methods for QTL mapping and draw an appropriate conclusion accordingly.

三種分析方式得到的結果非常相似，僅有些微差異，說明了即使由不同角度切入分析，由於統計分析方法都是正確的，都可以讓我們得到同樣正確的結論。
```{r}
plot(out.mr,out.em,out.hk, col=c(1,2,4))
legend("topright",c("MR","EM","HK"), col=c(1,2,4), lty=1)
```





