---
title: "Homework 15"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Compute the regression coefficients for fish = β0+β1*copepod and use randomization method to generate null distribution of β1 and test whether β1 is significantly different from null (with randomization 5000 times). Report your p-value.

#### loading package and data
```{r}
library(knitr)
library(ggplot2)

#### Read enviANDdensity.csv into R
enviANDdensity <- read.csv(file="~/Desktop/HW15_B04B01036/enviANDdensity.csv", header=TRUE)

# Because we only need the data in "FishDensity..ind..1000m3." and "CopepodDensity.ind..m3.", I take these two columns out and set as "FishDensity" and "CopepodDensity" separately.
FishDensity <- enviANDdensity$FishDensity..ind..1000m3.
CopepodDensity <- enviANDdensity$CopepodDensity.ind..m3.
```

#### Compute the β0 & β1 for 
#### fish(y) = β0 + β1 * copepod(x)
```{r}
# (same as HW3)
y  <- matrix(FishDensity, nrow=length(FishDensity), ncol=1, byrow = TRUE) # put the original sample into y
x  <- matrix(c(rep(1, times=length(FishDensity)),CopepodDensity), nrow=length(FishDensity), ncol=2, byrow = FALSE)  # put the original sample into x

β1 <- solve(t(x) %*% x) %*% t(x) %*% y
β0 <- y - x %*% β1  # fish(y) = β0 + β1 * copepod(x)
β1
```

#### Randomization method to generate null distribution of β1
#### null hypothesis H0: β1=0
#### randomization 5000 times
```{r}
randomization_times <- 5000
β1_null <- numeric()
for (i in 1:randomization_times) {
    shuffle = order(runif(length(FishDensity)))
    #c alculate β1
    Y = FishDensity[shuffle]
    X = cbind(1, CopepodDensity)
    b = solve(t(X) %*% X) %*% (t(X) %*% Y)
    β1_null = append(β1_null, b)
}
β1_null = t(matrix(β1_null, 2, randomization_times))
```

#### Report the p-value
```{r}
# p-value (two tailed test) 
# I use absolute function to make sure that I take both tails into consideration.

#calculate p-value for the two tailed test
p_value_1 = sum(abs(β1_null[, 2]) > abs(β1[2])) / randomization_times

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q1 = c(-β1[2], seq(min(β1_null[, 2]), max(β1_null[, 2]), length.out=5), β1[2])
p_q1 <- ggplot(as.data.frame(β1_null[, 2]), aes(x = β1_null[, 2])) +
        geom_density(fill = "#999999", col = "#333333", alpha = 0.5) +
        geom_vline(xintercept = c(-β1[2], β1[2]), col = "red") +
        labs(title = "Permutation", x = "β1") +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > β1[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -β1[2]), aes(x = x, y = y), col = "red",  fill = "red")

```

We can reject the null hypothesis because the p-value is smaller than 0.05.
Thus, we can say that β1 is significantly different from 0.
```{r}
p_value_1 # The p-value of the test on H0: B1=0.
```


#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 2. Randomization test whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper. (Assume all data are independent and use all 34 stations.) Report your p-value.

#### Import data
```{r}
copepod_composition <- read.table("~/Desktop/HW15_B04B01036/copepod_composition.txt", header = TRUE)
cop_density <- read.table("~/Desktop/HW15_B04B01036/cop_density.txt", header = TRUE)
copepodSPlist <- read.table("~/Desktop/HW15_B04B01036/copepodSPlist.txt")
```

#### Oncaea Venusta and Canthocalanus pauper
Because we only need the data in "Oncaea Venusta" and "Canthocalanus pauper", I take these two columns out and set as "Oncaea_Venusta" and "Canthocalanus_pauper" separately.
```{r}
# calculate copepod density for each species for each cruise station
copepod_composition <- copepod_composition / 100 # turn into percentage
copepod_composition_density_0 <- apply(copepod_composition, 1, function(x) x * cop_density[,1])
copepod_composition_density <- t(copepod_composition_density_0)

# Oncaea Venusta and Canthocalanus pauper
Oncaea_Venusta <- copepod_composition_density[  which (copepodSPlist[, 1] == "Oncaea_venusta"),  ]
Canthocalanus_pauper <- copepod_composition_density[  which (copepodSPlist[, 1] == "Canthocalanus_pauper") , ]
```

#### Randomization method to generate null distribution of θ1−θ2
#### null hypothesis H0: θ1−θ2=0
#### randomization 5000 times
```{r}
randomization_times <- 5000

difference <- numeric(randomization_times)
for(i in 1:randomization_times){
    randomization <- c(rep(c(1, 2), 34))[order(runif(68))]
    data <- cbind(c(Oncaea_Venusta, Canthocalanus_pauper), randomization)
    g1 <- subset(data, randomization == 1)
    g2 <- subset(data, randomization == 2)
    difference[i] <-  mean(g1) - mean(g2)
}
```

#### Plot the results
```{r}
#calculate estimated difference between Oncaea Venusta and Canthocalanus pauper density
est_difference <- mean(Oncaea_Venusta) - mean(Canthocalanus_pauper)

#calculate p-value
p_value_2 <- sum(abs(difference) > abs(est_difference)) / randomization_times

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q2 <- c(-est_difference, seq(min(difference), max(difference), length.out = 5), est_difference)
p_q2 <- ggplot(as.data.frame(difference), aes(x = difference)) +
        geom_density(fill = "#999999", col = "#333333", alpha = 0.5) +
        geom_vline(xintercept = c(-est_difference, est_difference), col = "red") +
        labs(title = "Permutattion", x = "theta1 - theta2") +
        scale_x_continuous(breaks = lab_q2, labels = format(lab_q2, digits = 2))

#add shaded area to indicate significant values
d_q2 <- ggplot_build(p_q2)$data[[1]]
p_q2 + geom_area(data = subset(d_q2, x > est_difference), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q2, x < -est_difference), aes(x = x, y = y), col = "red", fill = "red")
```

We can reject the null hypothesis because the p-value is smaller than 0.05.
Thus, we can say that significant difference does not exist between the density of Oncaea Venusta and Canthocalanus pauper.
```{r}
p_value_2 # The p-value of the test on H0:θ1−θ2=0.
```