---
title: "Homework 15"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Compute the regression coefficients for fish = β0+β1*copepod and use randomization method to generate null distribution of β1 and test whether β1 is significantly different from null (with randomization 5000 times). Report your p-value.

#### loading package and data
```{r}
library(knitr)
library(ggplot2)

#### Read enviANDdensity.csv into R
enviANDdensity <- read.csv(file="~/Desktop/HW15_B04B01036/enviANDdensity.csv", header=TRUE)

# Because we only need the data in "FishDensity..ind..1000m3." and "CopepodDensity.ind..m3.", I take these two columns out and set as "FishDensity" and "CopepodDensity" separately.
FishDensity <- enviANDdensity$FishDensity..ind..1000m3.
CopepodDensity <- enviANDdensity$CopepodDensity.ind..m3.
```

#### Compute the β0 & β1 for 
#### fish(y) = β0 + β1 * copepod(x)
```{r}
# (same as HW3)
y  <- matrix(FishDensity, nrow=length(FishDensity), ncol=1, byrow = TRUE) # put the original sample into y
x  <- matrix(c(rep(1, times=length(FishDensity)),CopepodDensity), nrow=length(FishDensity), ncol=2, byrow = FALSE)  # put the original sample into x

β1 <- solve(t(x) %*% x) %*% t(x) %*% y
β0 <- y - x %*% β1  # fish(y) = β0 + β1 * copepod(x)
β1
```

#### Randomization method to generate null distribution of β1
#### null hypothesis H0: β1=0
#### randomization 5000 times
```{r}
randomization_times = 5000
β1_null = NULL

for (i in 1:randomization_times) {
    # randomization
    RUNIF <- runif(n=length(FishDensity), min=0, max=1)
    randomization <- order(RUNIF)

    #define randomized dependent variable and independent variables and calculate coefficients
    y_random  <- matrix(FishDensity[randomization], nrow=length(FishDensity), ncol=1, byrow = TRUE) # put the randomized-ordered sample into y
    x_random  <- matrix(c(rep(1, times=length(FishDensity)),CopepodDensity), nrow=length(FishDensity), ncol=2, byrow = FALSE)  # put the original sample into x

    β1 <- solve(t(x_random) %*% x_random) %*% t(x_random) %*% y_random
    β1_null = append(β1_null, β1)
}

β1_null = t(matrix(β1_null, 2, randomization_times))
```

#### Report the p-value
```{r}
# p-value (two tailed test) 
# I use absolute function to make sure that I take both tails into consideration.
p_value_q1 = sum(abs(β1_null[, 2]) > abs(β1[2])) / randomization_times

# plot the results
lab_q1 = c(-β1[2], seq(min(β1_null[, 2]), max(β1_null[, 2]), length.out=5), β1[2])
p_q1 <- ggplot(as.data.frame(β1_null[, 2]), aes(x = β1_null[, 2])) +
        geom_density(fill = "#99CCFF", col = "#3399FF", alpha = 0.5) +
        geom_vline(xintercept = c(-β1[2], β1[2]), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(β1)[1]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(β1)[1])) +
        annotate("text", x = β1[2], y = 10, label = "Esimated~hat(β1)[1]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > β1[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -β1[2]), aes(x = x, y = y), col = "red",  fill = "red")
```

The p-value of our test on H0:β1=0 is 0.00009, therefore we reject the null hypothesis and conclude that β1 is significantly different from 0.


#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 2. Randomization test whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper. (Assume all data are independent and use all 34 stations.) Report your p- value.

Import data
```{r}
#import copepod data
species <- read.table("~/Desktop/HW15_B04B01036/copepod_composition.txt", header = TRUE)

#import density data
dens <- as.vector(read.table("~/Desktop/HW15_B04B01036/cop_density.txt", header = TRUE)[[1]])

#import species name
species_name <- read.table("~/Desktop/HW15_B04B01036/copepodSPlist.txt", sep = "\t")
```

Extract Oncaea Venusta and Canthocalanus pauper data
```{r}
#convert species frequency into percentage
species = species / 100

#calculate copepod density for each species for each cruise station
species.density = t(apply(species, 1, function(x) x * dens))

#extract the density of Oncaea Venusta and Canthocalanus pauper
OV = species.density[grep("Oncaea venusta", species_name[, 1]), ]
CP = species.density[grep("Canthocalanus pauper", species_name[, 1]), ]

#show partial data
kable(head(cbind(OV, CP)), digits = 2, col.names = c("Oncaea Venusta", "Canthocalanus pauper"), align = "l", caption = "Oncaea Venusta and Canthocalanus pauper density data (partial)")
```

Generate samples for permutation test for null hypothesis H0:θ1−θ2=0
```{r}
diff = numeric(randomization_times)
for(i in 1:randomization_times){
    randomization = c(rep(c(1, 2), 34))[order(runif(68))]
    data = cbind(c(OV, CP), randomization)
    g1 = subset(data, randomization == 1)
    g2 = subset(data, randomization == 2)
    diff[i] = mean(g1) - mean(g2)
}
```

Plot the distribution of permuted θ1^−θ2^
θ
1
θ
2
 samples and report p-value for the test
```{r}
#calculate estimated difference between Oncaea Venusta and Canthocalanus pauper density
est_diff = mean(OV) - mean(CP)

#calculate p-value
p_value_q2 = sum(abs(diff) > abs(est_diff)) / randomization_times

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q2 = c(-est_diff, seq(min(diff), max(diff), length.out = 5), est_diff)
p_q2 <- ggplot(as.data.frame(diff), aes(x = diff)) +
        geom_density(fill = "#99CCFF", col = "#3399FF", alpha = 0.5) +
        geom_vline(xintercept = c(-est_diff, est_diff), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(theta)[1]-hat(theta)[2]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(theta)[1]-hat(theta)[2])) +
        annotate("text", x = est_diff, y = 0.01, label = "Esimated~hat(theta)[1]-hat(theta)[2]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q2, labels = format(lab_q2, digits = 2))

#add shaded area to indicate significant values
d_q2 <- ggplot_build(p_q2)$data[[1]]
p_q2 + geom_area(data = subset(d_q2, x > est_diff), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q2, x < -est_diff), aes(x = x, y = y), col = "red", fill = "red")
```

The p-value of our test on H0:θ1−θ2=0 is 0.01011, therefore we reject the null hypothesis and conclude that significant difference does exist between the density of Oncaea Venusta and Canthocalanus pauper.