---
title: "Homework 15"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Compute the regression coefficients for fish = β0+β1*copepod and use randomization method to generate null distribution of β1 and test whether β1 is significantly different from null (with randomization 5000 times). Report your p-value.

#### loading package and data
```{r}
library(knitr)
library(ggplot2)

#### Read enviANDdensity.csv into R
enviANDdensity <- read.csv(file="~/Desktop/HW15_B04B01036/enviANDdensity.csv", header=TRUE)

# Because we only need the data in "FishDensity..ind..1000m3." and "CopepodDensity.ind..m3.", I take these two columns out and set as "FishDensity" and "CopepodDensity" separately.
FishDensity <- enviANDdensity$FishDensity..ind..1000m3.
CopepodDensity <- enviANDdensity$CopepodDensity.ind..m3.
```

#### Compute the β0 & β1 for 
#### fish(y) = β0 + β1 * copepod(x)
```{r}
# (same as HW3)
y  <- matrix(FishDensity, nrow=length(FishDensity), ncol=1, byrow = TRUE) # put the original sample into y
x  <- matrix(c(rep(1, times=length(FishDensity)),CopepodDensity), nrow=length(FishDensity), ncol=2, byrow = FALSE)  # put the original sample into x

β1 <- solve(t(x) %*% x) %*% t(x) %*% y
β0 <- y - x %*% β1  # fish(y) = β0 + β1 * copepod(x)
β1
```

#### Randomization method to generate null distribution of β1
#### null hypothesis H0: β1=0
#### randomization 5000 times
```{r}
randomization_times <- 5000
β1_null <- NULL

for (i in 1:randomization_times) {
    # randomization
    RUNIF <- runif(n=length(FishDensity), min=0, max=1)
    randomization <- order(RUNIF)

    # compute β1
    y_random  <- FishDensity[randomization]  # put the randomized-ordered sample into y
    x_random  <- matrix(c(rep(1, times=length(FishDensity)),CopepodDensity), nrow=length(FishDensity), ncol=2, byrow = FALSE)  # put the original sample into x

    β1 <- solve(t(x_random) %*% x_random) %*% t(x_random) %*% y_random
    β1_null = append(β1_null, β1)
}

β1_null = matrix(β1_null, randomization_times ,2, byrow = TRUE)

β1_null


resamp_n = 50000
size = length(FishDensity)
beta_sample = numeric()
for (i in 1:resamp_n) {
    #create a lish of indices for randomization
    shuffle = order(runif(size))
    
    #define randomized dependent variable and independent variables and calculate coefficients
    Y = FishDensity[shuffle]
    X = cbind(1, CopepodDensity)
    b = solve(t(X) %*% X) %*% (t(X) %*% Y)
    beta_sample = append(beta_sample, b)
}
beta_sample = t(matrix(beta_sample, 2, resamp_n))
beta_sample

β1_null = beta_sample
```






#### Report the p-value
```{r}
# p-value (two tailed test) 
# I use absolute function to make sure that I take both tails into consideration.
p_value_q1 <- sum(abs(β1_null[, 2]) > abs(β1[2])) / randomization_times

# plot the results
lab_q1 <- c(-β1[2], seq(min(β1_null[, 2]), max(β1_null[, 2]), length.out=5), β1[2])
p_q1 <- ggplot(as.data.frame(β1_null[, 2]), aes(x = β1_null[, 2])) +
        geom_density(fill = "#999999", col = "#333333", alpha = 0.5) +
        geom_vline(xintercept = c(-β1[2], β1[2]), col = "red") +
        labs(title = "Null hypothesis distribution (5000 times)", x = "Randonmized β1") +
        annotate("text", x = β1[2], y = 10, label = "Esimated~hat(β1)[1]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > β1[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -β1[2]), aes(x = x, y = y), col = "red",  fill = "red")


###

#calculate p-value for the two tailed test
p_value_q1 = sum(abs(beta_sample[, 2]) > abs(beta[2])) / resamp_n

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q1 = c(-beta[2], seq(min(beta_sample[, 2]), max(beta_sample[, 2]), length.out=5), beta[2])
p_q1 <- ggplot(as.data.frame(beta_sample[, 2]), aes(x = beta_sample[, 2])) +
        geom_density(fill = "#99CCFF", col = "#3399FF", alpha = 0.5) +
        geom_vline(xintercept = c(-beta[2], beta[2]), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(beta)[1]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(beta)[1])) +
        annotate("text", x = beta[2], y = 10, label = "Esimated~hat(beta)[1]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > beta[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -beta[2]), aes(x = x, y = y), col = "red",  fill = "red")


###

#calculate p-value for the two tailed test
p_value_q1 = sum(abs(beta_sample[, 2]) > abs(beta[2])) / resamp_n

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q1 = c(-beta[2], seq(min(beta_sample[, 2]), max(beta_sample[, 2]), length.out=5), beta[2])
p_q1 <- ggplot(as.data.frame(beta_sample[, 2]), aes(x = beta_sample[, 2])) +
        geom_density(fill = "#99CCFF", col = "#3399FF", alpha = 0.5) +
        geom_vline(xintercept = c(-beta[2], beta[2]), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(beta)[1]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(beta)[1])) +
        annotate("text", x = beta[2], y = 10, label = "Esimated~hat(beta)[1]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > beta[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -beta[2]), aes(x = x, y = y), col = "red",  fill = "red")




```
β1_null beta_sample / β1 beta / randomization_times resamp_n


```{r}

#calculate p-value for the two tailed test
p_value_q1 = sum(abs(beta_sample[, 2]) > abs(beta[2])) / resamp_n

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q1 = c(-beta[2], seq(min(beta_sample[, 2]), max(beta_sample[, 2]), length.out=5), beta[2])
p_q1 <- ggplot(as.data.frame(beta_sample[, 2]), aes(x = beta_sample[, 2])) +
        geom_density(fill = "#99CCFF", col = "#3399FF", alpha = 0.5) +
        geom_vline(xintercept = c(-beta[2], beta[2]), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(beta)[1]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(beta)[1])) +
        annotate("text", x = beta[2], y = 10, label = "Esimated~hat(beta)[1]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q1, labels = format(lab_q1, digits = 1))

#add shaded area to indicate significant values
d_q1 <- ggplot_build(p_q1)$data[[1]]
p_q1 + geom_area(data = subset(d_q1, x > beta[2]), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q1, x < -beta[2]), aes(x = x, y = y), col = "red",  fill = "red")
```

The p-value of our test on H0:β1=0 is 0.00009, therefore we reject the null hypothesis and conclude that β1 is significantly different from 0.


#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 2. Randomization test whether significant difference exists between the density of Oncaea Venusta and Canthocalanus pauper. (Assume all data are independent and use all 34 stations.) Report your p-value.

#### Import data
```{r}
copepod_composition <- read.table("~/Desktop/HW15_B04B01036/copepod_composition.txt", header = TRUE)
cop_density <- read.table("~/Desktop/HW15_B04B01036/cop_density.txt", header = TRUE)
copepodSPlist <- read.table("~/Desktop/HW15_B04B01036/copepodSPlist.txt")
```

#### Oncaea Venusta and Canthocalanus pauper
Because we only need the data in "Oncaea Venusta" and "Canthocalanus pauper", I take these two columns out and set as "Oncaea_Venusta" and "Canthocalanus_pauper" separately.
```{r}
# calculate copepod density for each species for each cruise station
copepod_composition <- copepod_composition / 100 # turn into percentage
copepod_composition_density <- (apply(copepod_composition, 1, function(x) x * cop_density[,1]))
copepod_composition_density <- t(copepod_composition_density)

# Oncaea Venusta and Canthocalanus pauper
Oncaea_Venusta <- copepod_composition_density[  ,which (copepodSPlist[, 1] == "Oncaea_venusta")  ]
Canthocalanus_pauper <- copepod_composition_density[  ,which (copepodSPlist[, 1] == "Canthocalanus_pauper")  ]
```

#### Randomization method to generate null distribution of θ1−θ2
#### null hypothesis H0: θ1−θ2=0
#### randomization 5000 times
```{r}
randomization_times <- 5000

difference <- numeric(randomization_times)
for(i in 1:randomization_times){
    randomization <- c(rep(c(1, 2), 34))[order(runif(68))]
    data <- cbind(c(Oncaea_Venusta, Canthocalanus_pauper), randomization)
    g1 <- subset(data, randomization == 1)
    g2 <- subset(data, randomization == 2)
    difference[i] <-  mean(g1) - mean(g2)
}
```

#### Plot the results
```{r}
#calculate estimated difference between Oncaea Venusta and Canthocalanus pauper density
est_difference <- mean(Oncaea_Venusta) - mean(Canthocalanus_pauper)

#calculate p-value
p_value_q2 <- sum(abs(difference) > abs(est_difference)) / randomization_times

#create density plot for the permuted samples and add vertical lines at the estimated value
lab_q2 <- c(-est_difference, seq(min(difference), max(difference), length.out = 5), est_difference)
p_q2 <- ggplot(as.data.frame(difference), aes(x = difference)) +
        geom_density(fill = "#999999", col = "#333333", alpha = 0.5) +
        geom_vline(xintercept = c(-est_difference, est_difference), col = "red") +
        labs(title = expression(Distribution~of~permuted~hat(theta)[1]-hat(theta)[2]~samples~"for"~null~hypothesis), x = expression(Permuted~hat(theta)[1]-hat(theta)[2])) +
        annotate("text", x = est_difference, y = 0.01, label = "Esimated~hat(theta)[1]-hat(theta)[2]", hjust = 1.1, parse = T) +
        scale_x_continuous(breaks = lab_q2, labels = format(lab_q2, digits = 2))

#add shaded area to indicate significant values
d_q2 <- ggplot_build(p_q2)$data[[1]]
p_q2 + geom_area(data = subset(d_q2, x > est_difference), aes(x = x, y = y), col = "red", fill = "red") +
    geom_area(data = subset(d_q2, x < -est_difference), aes(x = x, y = y), col = "red", fill = "red")
```

The p-value of our test on H0:θ1−θ2=0 is 0.01011, therefore we reject the null hypothesis and conclude that significant difference does exist between the density of Oncaea Venusta and Canthocalanus pauper.