---
title: "Homework 14"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## Assume t0=0 (two parameters). Please estimate the posterior distributions of parameters using the method Markov chain Monte Carlo (MCMC) for the average length-at-age for the female Pacific Hake (obs. sigma=0.032), considering a multiplicative error model with log-normal distribution.

#### von Bertalanffy Growth Function (Considering a log-normal distribution)
#### L(t) = L∞ (1−exp(−k(t−t0))) * e^ε   ,   ε ~ N(0,σ^2)

#### L∞ : the hypothesized (mean) maximum length
####  K : growth rate
####  t : current age
#### t0 : the hypothesized age when body length is 0

#### Prior distribu>ons
####  Linf ~ U [ 40, 100]
####     K ~ U [0.1, 0.6]

## loading package and data
```{r}
library(knitr)
library(ggplot2)

Age = c(1,2,3.3,4.3,5.3,6.3,7.3,8.3,9.3,10.3,11.3,12.3,13.3)
Length = c(15.4,28.03,41.18,46.2,48.23,50.26,51.82,54.27,56.98,58.93,59,60.91,61.83)

data = cbind(Age, Length)

kable(data, col.names = c("Age (year)", "Female mean length (cm)"))
```

## Check the plot of [ Length ~ Age ]
```{r}
dataset = data.frame(Age, Length) # combine the data
colnames(dataset) <- c("Age", "Length") # rename the columns for plotting
ggplot(dataset, aes(x = Age, y = Length)) +
    geom_point()
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## likelihood section
```{r}
VBGF <-function(x, Linf, K){
    y = Linf * (1 - exp(- K * (x - 0)))
    return(y)
}


lognormal_like = function(Linf, K){
  
  Like = numeric(length(Length))
  neg_likelihood = numeric(length(Length))
  ypred = VBGF(Age, Linf, K)
  Dev2 =(log(Length) - log(ypred)) ^ 2
  sigma = 0.032
  
  for (i in 1:length(Length)){
    Like[i] = (1 / (Age[i] * sqrt(2 * pi) * sigma)) * exp(-Dev2[i] / (2 * sigma ^ 2))
    neg_likelihood[i] = -log(Like[i])
  }
  
  totoal_Like = exp(-sum(neg_likelihood))
  
  Outs <- NULL
  Outs$totoal_Like <- totoal_Like
  Outs$neg_likelihood <- sum(neg_likelihood)
  Outs$ypred <- ypred
  Outs$Linf <- Linf
  Outs$K <- K
  
  return(Outs) # But actually, I will only use "Outs$neg_likelihood".
}
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## MCMC section
```{r}

DoMCMC<-function(Xinit, Ndim, Nsim = 1000, Nburn = 0, Nthin = 1)
  {  
  # jump function steps
  Linf_jump_max = 100
  Linf_jump_min = -100
  K_jump_max = 0.05
  K_jump_min = -0.05
  
  Xcurr <- Xinit
  Fcurr <- -1 * lognormal_like(Linf = Xcurr[1], K = Xcurr[2])$neg_likelihood
  Outs2 <- matrix(-9999, nrow = (Nsim - Nburn), ncol = (Ndim + 1))
  Ipnt <- 0; Icnt <- 0
  
  
  for(Isim in 1:Nsim)
    { 
    # jump function
    Xnext = NULL
        
    repeat{ # Linf prior: Linf ~ U [40, 100]
      Xnext[1] = Xcurr[1] + runif(1, 0, 1) * (Linf_jump_max - Linf_jump_min) + Linf_jump_min
      if(100>=Xnext[1] && Xnext[1]>=40){
        break
        }
    }
    
    repeat{ # K prior: K ~ U [0.1, 0.6]
      Xnext[2] = Xcurr[2] + runif(1, 0, 1) * (K_jump_max - K_jump_min) + K_jump_min
      if(0.6>=Xnext[2] && Xnext[2]>=0.1){
        break
        }
    }
    

    # evaluate the ratio
    Fnext <- -1 * lognormal_like(Linf = Xnext[1], K = Xnext[2])$neg_likelihood
     Rand1 <- log(runif(1,0,1))

     if (Fnext > Fcurr+Rand1)
       {Fcurr <- Fnext; Xcurr <- Xnext }  

    # thinning and burn in 
    if(Isim %% Nthin == 0){  # thinning
      Ipnt <- Ipnt + 1
      if (Ipnt > Nburn){ # burn in
        Icnt <- Icnt + 1
        Outs2[Icnt, ] <- c(Xcurr, Fcurr)
        }
      }

  }

  
  return(Outs2[1:Icnt,])
}

```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## Posterior section
```{r}
Posterior_result <- DoMCMC(Xinit = c(100, 0.1), Ndim = 2, Nsim = 50000, Nburn = 100, Nthin = 10)
colnames(Posterior_result) <- c("Linf", "K", "neg_likelihood")
head(Posterior_result)

par(mfrow=c(2,2),mar=c(5,5,5,2))

plot(Posterior_result[,1],main="",xlab="Linf",type="o")
plot(Posterior_result[,2],main="",xlab="K",type="o")

hist(Posterior_result[,1],
     main="",
     xlab="Linf",
     col="gray", 
     breaks = seq(min(Posterior_result[,1]), max(Posterior_result[,1]), length.out = 100))
abline(v=median(Posterior_result[,1]), col ="red", lwd = 5)

hist(Posterior_result[,2],
     main="",
     xlab="K",
     col="gray",
     breaks = seq(min(Posterior_result[,2]), max(Posterior_result[,2]), length.out = 1000))
abline(v=median(Posterior_result[,2]), col ="red", lwd = 5)

print(paste("Median of Linf = ",round(median(Posterior_result[,1]),4),sep=""))
print(paste("Median of K = ",round(median(Posterior_result[,2]),4),sep=""))
```


## Compare MCMC with MLE

#### When we used the MLE in previous homework, we got:
#### Linf = 61.08735457
#### K = 0.30279632
#### Both parameters are quite similar as the results in this homework because
#### 1) in prior distribution we used the uniform distribution and
#### 2) we used sigma = 0.032 to start, which is the same value of MLE results.



















