---
title: "Homework 13"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

The von Bertalanffy growth function is used to describe body length of fish as a function of its age. It is defined as L(t)=L∞(1−e−K(t−t0)) where L∞ is the hypothesized (mean) maximum length, K is growth rate, t is current age and t0 is the hypothesized age when body length is 0.

In this assignment, we try to model body length of fish using the following regression model: L(t)=L∞(1−e−K(t−t0))∗eϵ with ϵ ~ N(0,σ2). Parameters will be estimated using the Bayesian grid search with priors L∞ ~ U(40,100) and K ~ U(0.1,0.6).

#### Assume t0 = 0 and sigma=0.032, please estimate the parameters of L∞ and K by using the method of Bayesian Grid Search for the average length-at-age data Pacific Hake, respec%vely, considering a mul%plica%ve error model with log-normal distribu%on with the prior informa%on of K~U[0.1,0.6] and L∞~U[40,100].

Q. Assume t0=0 (two parameters). Please estimate the posterior distributions of parameters using the method Bayesian grid search for the average length-at-age for the female Pacific Hake, respectively, considering a multiplicative error model with log-normal distribution.

## Models
#### 1 von Bertalanffy Growth Function
#### -> L(t) = L∞ (1−exp(−k(t−t0)))
#### 2 Gompertz growth model
#### -> L(t) = L∞ exp(-exp(-k2(t-((ln(λ)-ln(k2))/k2 ))))
#### 3 Logistic model 
#### -> L(t) = L∞ (1 + exp(-k3(t-t3)))^(-1)
#### 4 Schnute–Richards model] Bonus
#### -> L(t) = L∞ exp(1 + δ*e^(-k4 t^v)) ^ (1/γ)

## Considering a log-normal distribution
#### L(t) = Model * e^ε   ,   ε ~ N(0,σ^2)

## loading package and data
```{r}
library(knitr)
library(ggplot2)

Age = c(1,2,3.3,4.3,5.3,6.3,7.3,8.3,9.3,10.3,11.3,12.3,13.3)
Length = c(15.4,28.03,41.18,46.2,48.23,50.26,51.82,54.27,56.98,58.93,59,60.91,61.83)

data = cbind(Age, Length)

kable(data, col.names = c("Age (year)", "Female mean length (cm)"))
```

## Check the plot of [ Length ~ Age ]
```{r}
dataset = data.frame(Age, Length) # combine the data
colnames(dataset) <- c("Age", "Length") # rename the columns for plotting
ggplot(dataset, aes(x = Age, y = Length)) +
    geom_point()
```


#### --------------------------------------------------------------------------------------------------------------------------------------------------------
```{r}

 # initial parameters 
 obs_N=NULL
 sigma=NULL
 
 obs_N[9] = 453
 obs_N[16] = 559
 obs_N[26] = 2280

 sigma[9] = 0.4
 sigma[16] = 0.47
 sigma[26] = 0.36
```

# (prof)
## likelihood section

```{r}
 # population dynamics model
 pop_model <- function(r,N0){
  
 N = NULL
 Dev2 = NULL
 N[1] = N0
 for(t in 2:33){
 N[t] = N[t-1]*(1+r)  
 Dev2[t]=(log(obs_N[t]/N[t]))^2
 }

 Like_1 = 1/obs_N[9]*(1/sqrt(2*pi*sigma[9]*sigma[9]))*exp(-Dev2[9]/(2*sigma[9]*sigma[9]))
 Like_2 = 1/obs_N[16]*(1/sqrt(2*pi*sigma[16] *sigma[16]))*exp(-Dev2[16]/(2*sigma[16]*sigma[16]))
 Like_3 = 1/obs_N[26]*(1/sqrt(2*pi*sigma[26] *sigma[26]))*exp(-Dev2[26]/(2*sigma[26]*sigma[26]))
 neg_likelihood =  -(log(Like_1)+log(Like_2)+log(Like_3))
 totoal_like = exp(-neg_likelihood)
 return(totoal_like)
 }
 #------------------------------------------------------------------------------- 
 pop_model(r=0.1,N0=100)
```

VBGF <-function(x, Linf, K){
    y = Linf * (1 - exp(- K * (x - 0)))
    return(y)
}

lognormal_like = function(Linf, K){
    
    like = numeric(length(length))
    LL = numeric(length(length))
    ypred = VBGF(age, Linf, K)
    dev2 =(log(length) - log(ypred)) ^ 2
    sigma = 0.2 # 0.032
    
    for (i in 1:length(length)){
        like[i] = (1 / (age[i] * sqrt(2 * pi) * sigma)) * exp(-dev2[i] / (2 * sigma ^ 2))
        LL[i] = log(like[i])
    }
    
    tot_like = exp(sum(LL))
    return(tot_like)
}


#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## Prior section
(grid search)
```{r}
# define the grid of each parameter
Linf_grid = seq(40, 100, length.out = 1000) # L∞ ~ U(40,100)  # 1000??????? 0.1
K_grid = seq(0.1, 0.6, length.out = 1000) # K ~ U(0.1,0.6) # 0.001?

# prior porbability
pLinf_grid = rep(1, length(Linf_grid))
pK_grid = rep(1, length(K_grid))
pTheta_matrix = expand.grid(Linf = pLinf_grid, K = pK_grid)

# expand the combination of grid 
Theta_matrix = expand.grid(Linf = Linf_grid, K = K_grid)
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

## Posterior section
```{r}
# negtive likelihood values of each grid
pDataGivenTheta = mapply(lognormal_like, Theta_matrix[,1], Theta_matrix[,2]) # pop_model -> lognormal_like # likelihood

# Theta_matrix "Linf" "K"

# Compute the posterior density
pData = sum(pDataGivenTheta * pTheta_matrix[, 1] * pTheta_matrix[, 2])
pThetaGivenData = (pDataGivenTheta * pTheta_matrix[, 1] * pTheta_matrix[, 2]) / pData

# arrange the results
RESULTs = cbind(Theta_matrix, pThetaGivenData)
```

# Plot posterior distributions of parameters L∞ and K  @@@@@@@@
```{r}
par(mfrow=c(2,1),mar=c(4,4,2,2))

# calculate the marginal probability
marginal_Linf = with(RESULTs, tapply(pThetaGivenData, Linf, sum))
plot(Linf_grid, marginal_Linf, type = "h", lwd = 3, xlab = expression(L[infinity]), ylab = bquote(paste("p(", theta, "|D)"))) ######@@@@@@ xlab ylab xlab="r",ylab=bquote(paste("p(",theta,"|D)"))) ?????????


# calculate the marginal probability
marginal_K = with(RESULTs, tapply(pThetaGivenData, K, sum))
plot(K_grid, marginal_K, type = "h", lwd = 3, xlab = "K", ylab = bquote(paste("p(", theta, "|D)"))) ############# xlab="N1973",ylab=bquote(paste("p(",theta,"|D)"))) 

abline =  (v=linf[which.max(marginal_K),col = "red", lwd=2]) # median mean dont use max
  text
  

# compare w/ ML / cuz prior -> uniform / same sigma start

```






















