---
title: "Homework 12"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Please estimate the model parameters of four different growth functions (von Bertalanffy; Gompertz; Logis>c; Schnute–Richards) using the method maximum likelihood for the length-at-age for female Pacific hake (Merluccius productus) considering a log-normal distribution

## Models
#### [ 1 von Bertalanffy Growth Function ]
#### -> L(t) = L∞ (1−exp(−k(t−t0)))
#### [ 2 Gompertz growth model ]
#### -> L(t) = L∞ exp(-exp(-k2(t-((ln(λ)-ln(k2))/k2 ))))
#### [ 3 Logistic model ] 
#### -> L(t) = L∞ (1 + exp(-k3(t-t3)))^(-1)
#### [ 4 Schnute–Richards model] Bonus
#### -> L(t) = L∞ exp(1 + δ*e^(-k4 t^v)) ^ (1/γ)

## Considering a log-normal distribution
## L(t) = Model * e^ε   ,   ε ~ N(0,σ^2)

## loading package and data
```{r}
library(knitr)
library(ggplot2)

Age = c(1,2,3.3,4.3,5.3,6.3,7.3,8.3,9.3,10.3,11.3,12.3,13.3)
Length = c(15.4,28.03,41.18,46.2,48.23,50.26,51.82,54.27,56.98,58.93,59,60.91,61.83)

data = cbind(Age, Length)

kable(data, col.names = c("Age (year)", "Female mean length (cm)"))
```

## Check the plot of [ Length ~ Age ]
```{r}
dataset = data.frame(Age, Length) # combine the data
colnames(dataset) <- c("Age", "Length") # rename the columns for plotting
ggplot(dataset, aes(x = Age, y = Length)) +
    geom_point()
```


## [ 1 von Bertalanffy Growth Function ]
## L(t) = L∞ (1−exp(−k(t−t0)))
#### L∞ = Linf = hypothesized (mean) maximum length
#### k  = growth rate
#### t  = current age
#### t0 = hypothesized age when body length is 0
```{r}
VBGF <- function(x, Linf = theta[1], k = theta[2], t0 = theta[3], theta = c(Linf, k, t0)){
    y = Linf * (1 - exp(-k * (x - t0)))
    return(y)
}
```

## [ 2 Gompertz growth model ]
## L(t) = L∞ exp(-exp(-k2(t-((ln(λ)-ln(k2))/k2 ))))
#### L∞ = Linf = hypothesized (mean) maximum length
#### k2 = rate of exponential decrease of the relative growth rate with age
#### t  = current age
#### λ  = theoretical initial relative growth rate at zero age
```{r}
GGM <- function(x, Linf = theta[1], k2 = theta[2], λ = theta[3], theta = c(Linf, k2, λ)){
    y = Linf * exp(- exp(- k2 * (x - (log(λ) - log(k2)) / k2)))
    return(y)
}
```

## [ 3 Logistic model ] 
## L(t) = L∞ (1 + exp(-k3(t-t3)))^(-1)
#### L∞ = Linf = hypothesized (mean) maximum length
#### k3 = relative growth rate parameter (with units year^-1)
#### t  = current age
#### t3 = corresponds to the inflection point of the sigmoidal curve
```{r}
LM <- function(x, Linf = theta[1], k3 = theta[2], t3 = theta[3], theta = c(Linf, k3, t3)){
    y = Linf * (1 + exp(-k3*(t-t3))) ^ (-1)
    return(y)
}
```

## [ 4 Schnute–Richards model ] Bonus
## L(t) = L∞ exp(1 + δ*e^(-k4 t^v)) ^ (1/γ)
#### L∞ = Linf = hypothesized (mean) maximum length
#### dimensionless parameters: δ, ν, and γ
#### k4 has units yr^(−ν)
```{r}
SRM <- function(x, Linf = theta[1], k4 = theta[2], δ = theta[3], ν = theta[4], γ = theta[5], theta = c(Linf, k4, δ, ν, γ)){
    y = Linf * exp(1 + δ*e^(-k4*(t^v))) ^ (1/γ)
    return(y)
}
```

## Log-likelihood function
```{r}
lognormal_like = function(x, y, FUN, Out, ...){  
    
    FUN = match.fun(FUN) ######
    Like = numeric(length(y))
    NLL = numeric(length(y))
    ypred = FUN(x, ...)
    Dev2 = (log(Length/ypred)) ^ 2
    sigma = sqrt(mean(Dev2))
    
    for (i in 1:length(Age)){ 
        Like[i] = (1 / (x[i] * sqrt(2 * pi) * sigma)) * exp(-Dev2[i] / (2 * sigma ^ 2))
        NLL[i] = -log(Like[i])
    }
    
    if(Out==T) cat(sum(NLL),Linf,K,t0,"\n")
    
        Outs <- NULL
        Outs$Pred <- ypred
        Outs$Length <- Length
        Outs$Dev2 <- Dev2
        Outs$sigma <- sigma
        Outs$Like <- Like
        Outs$NLL <- NLL 
        Outs$Obj <- sum(NLL)
        
    return(Outs)
}
```

## Objective function
```{r}
obj_function = function(theda, ...){ 
    Outs <- lognormal_like(Linf = theda[1], k = theda[2], t0 = theda[3], ...)
    obj <- Outs$Obj 
    return(obj)
}
```

## MLE of parameters
```{r}
model_VBGF = optim(c(60, 0.7, 0), obj_function, FUN = VBGF, x = Age, y = Length, Out = F, hessian = T)
model_GGM = optim(c(70, 0.6, 0), obj_function, FUN = GGM, x = Age, y = Length, Out = F, hessian = T)
model_LM = optim(c(59, 0.7, 0), obj_function, FUN = LM, x = Age, y = Length, Out = F, hessian = T)
model_SRM = optim(c(61, 0.7, 0), obj_function, FUN = SRM, x = Age, y = Length, Out = F, hessian = T)
```

## Results
```{r}
param_VBGF <- model_VBGF$par
param_GGM <- model_GGM$par
param_LM <- model_LM$par
param_SRM <- model_SRM$par

sd_VBGF <- sqrt(diag(solve(model_VBGF$hessian)))
sd_GGM <- sqrt(diag(solve(model_GGM$hessian)))
sd_LM <- sqrt(diag(solve(model_LM$hessian)))
sd_SRM <- sqrt(diag(solve(model_SRM$hessian)))

model_result_VBGF <- cbind(param_VBGF, sd_VBGF)
model_result_GGM <- cbind(param_GGM, sd_GGM)
model_result_LM <- cbind(param_LM, sd_LM)
model_result_SRM <- cbind(param_SRM, sd_SRM)

rownames(model_result_VBGF) <- c("L inf", "k", "t0")
rownames(model_result_GGM) <- c("L inf", "k2", "λ")
rownames(model_result_LM) <- c("L inf", "k3", "t3")
# rownames(model_result_SRM) <- c("L inf", "K4", "δ", "ν", "γ" ) ###???????

kable(model_result_VBGF, col.names = c("Estimated parameters", "Standard deviation"), caption = "1 von Bertalanffy growth function", digits = 3)
kable(model_result_GGM, col.names = c("Estimated parameters", "Standard deviation"), caption = "2 Gompertz growth model", digits = 3)
kable(model_result_LM, col.names = c("Estimated parameters", "Standard deviation"), caption = "3 Logistic model", digits = 3)
kable(model_result_SRM, col.names = c("Estimated parameters", "Standard deviation"), caption = "4 Schnute–Richards model", digits = 3)
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------
#### 2. Calculate the 95% confidence intervals for Linf (Professor said we don't need to do this on each parameter) of the growth functions by using the likelihood profile method

## Negative log-likelihood
```{r}
VBGF_like = lognormal_like(data[, 1], data[, 2], VBGF, theta = param_VBGF, Out = F)$Obj
GGM_like = lognormal_like(data[, 1], data[, 2], GGM, theta = param_GGM, Out = F)$Obj
LM_like = lognormal_like(data[, 1], data[, 2], LM, theta = param_LM, Out = F)$Obj
SRM_like = lognormal_like(data[, 1], data[, 2], SRM, theta = param_SRM, Out = F)$Obj
```

## 1 VBGF [ L∞ - 95% confidence interval ]
```{r}
obj_fun = function(theta, ...){ 
    Outs <- lognormal_like(K = theta[1], t0 = theta[2], ...)
    obj <- Outs$Obj
    return(obj)
}

Linf_like = numeric()
Linf_range = seq(from = param_VBGF[1] - 40, to = param_VBGF[1] + 100, length.out = 100)
for(i in 1:100){
        Linf_est = optim(param_VBGF[c(2, 3)], obj_fun, FUN = VBGF, Linf = Linf_range[i], x = data[, 1], y = data[ ,2], Out = F)$par
        Linf_like[i] = 2 * (- VBGF_like + lognormal_like(data[, 1], data[, 2], VBGF, Linf =  Linf_range[i], k = Linf_est[1], t0 = Linf_est[2], Out = F)$Obj)
}

Linf_accept = which(Linf_like < qchisq(0.95, 1))
Linf_ci = Linf_range[c(min(Linf_accept), max(Linf_accept))]
```

## 2 GGM [ L∞ - 95% confidence interval ]
```{r}
obj_fun = function(theta, ...){ 
    Outs <- lognormal_like(K2 = theta[1], lambda = theta[2], ...)
    obj <- Outs$Obj
    return(obj)
}

LGomp_like = numeric()
LGomp_range = seq(from = param_Gomp[1] - 40, to = param_Gomp[1] + 80, length.out = 100)
for(i in 1:100){
        LGomp_est = optim(param_Gomp[c(2, 3)], obj_fun, FUN = Gomp, Linf = LGomp_range[i], x = data[, 1], y = data[ ,2], Out = F)$par
        LGomp_like[i] = 2 * (- Gomp_like + lognormal_like(data[, 1], data[, 2], Gomp, Linf =  LGomp_range[i], K2 = LGomp_est[1], lambda = LGomp_est[2], Out = F)$Obj)
}

LGomp_accept = which(LGomp_like < qchisq(0.95, 1))
LGomp_ci = LGomp_range[c(min(LGomp_accept), max(LGomp_accept))]
```

## 3 LM [ L∞ - 95% confidence interval ]
```{r}
```

## 4 SRM [ L∞ - 95% confidence interval ]
```{r}
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------
#### 3. Conduct the model selection based on the AICc and quantify the plausibility of each model by using “Akaike weight (wi)”
```{r}
#calculate AICc for both models
aicc_VBGF = 2 * VBGF_like + 2 * 3 + (2 * 3 * 4) / (204 - 3 - 1)
aicc_GGM = 2 * GGM_like + 2 * 3 + (2 * 3 * 4) / (204 - 3 - 1)
aicc_LM = 2 * LM_like + 2 * 3 + (2 * 3 * 4) / (204 - 3 - 1)
aicc_SRM = 2 * LM_like + 2 * 3 + (2 * 3 * 4) / (204 - 3 - 1)

#find the best model with the lower AICc
min_aicc = min(aicc_VBGF, aicc_GGM, aicc_LM, aicc_SRM)

#calculate weights based on AICc
aicc_change = c(aicc_VBGF, aicc_GGM, aicc_LM, aicc_SRM) - min_aicc
weight = exp(-0.5 * aicc_change) / sum(exp(-0.5 * aicc_change))

#show results of AICc and weights
m_criterion = cbind(c(aicc_VBGF, aicc_GGM, aicc_LM, aicc_SRM), aicc_change, weight)
rownames(m_criterion) <- c("von Bertalanffy", "Gompertz", "Logistic", "Schnute–Richards")
colnames(m_criterion) <- c("AICc", "AICc change", "Weight")
kable(m_criterion)
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------
#### 4. Estimate the average model based on wi and plot all the growth curves together

## Average model based on wi
```{r}
pred_VBGF = VBGF(seq(from = min(data[ ,1]), to = max(data[, 1]), length.out = 100), theta = param_VBGF)
pred_GGM = GGM(seq(from = min(data[ ,1]), to = max(data[, 1]), length.out = 100), theta = param_GGM)
pred_LM = LM(seq(from = min(data[ ,1]), to = max(data[, 1]), length.out = 100), theta = param_LM)
pred_SRM = SRM(seq(from = min(data[ ,1]), to = max(data[, 1]), length.out = 100), theta = param_SRM)
pred_ave = weight[1] * pred_VBGF + weight[2] * pred_GGM + weight[3] * pred_LM + weight[4] * pred_SRM
```

## Plot all the growth curves together
#### ???????? 不用重複100 次？？？？？
```{r}
#prepare data for plotting
dataset = data.frame(rep(seq(from = min(data[ ,1]), to = max(data[, 1]), length.out = 100), 3), c(pred_VBGF, pred_GGM, pred_ave), c(rep("VBGF", 100), rep("Gompertz",100), rep("Average", 100)))
colnames(dataset) <- c("Age", "Esimated length", "Model")

#draw graph
p <- ggplot() +
    geom_line(aes(x = dataset$Age, y = dataset$`Esimated length`, col = dataset$Model)) +
    scale_color_discrete(name = "Model") +
    labs(title = "Regression model using von Bertalanffy and Gompertz growth function", x = "Age", y = "Fish length")
p + geom_point(aes(x = data[, 1], y = data[, 2]))

```




###@@@
```{r}
# MLE estimation and Experiment data
dataset = data.frame(Age, Length, VBGF(Age, pars[1], pars[2], pars[3]))
colnames(dataset) <- c("Age", "Actual length", "Esimated length")

ggplot(dataset, aes(x = Age)) +
    geom_point(aes(y = `Actual length`)) +
    geom_line(aes(y = `Esimated length`)) +
    labs(title = "Regression model using von Bertalanffy growth function", y = "Fish length")
```






