---
title: "Homework 10"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Use the “dominant” copepod species data (from HW1) and environmental data (from HW3). Apply constrained ordination: 
#### 1) determine an appropriate response model; 
#### 2) select a set of independent variables; 
#### 3) conduct ordination (RDA or CCA); 
#### 4) interpret the results. Try different plotting methods; e.g., to display your cluster analysis results on the triplot and interpret how the groups vary with respect to environmental variables.

#### Retrieve the dominant species from HW1
```{r}
enviANDdensity <- read.csv("~/R/enviANDdensity.csv", row.names = 1, header=T)

dominant = read.csv("~/R/dominant.csv",row.names = 1,header=T) # examples HW1
data = t(dominant)
data = data/sum(data) # get the propotion data

y = data
x = enviANDdensity
```

# RDA 
```{r}
library(vegan)

# normalize the dominant species data
y.log = log(y+1)    #log-transform y
y.chord = data.frame(matrix(0,nrow(y.log),ncol(y.log)))    #row normalization
rsum = apply(y.log,1,sum)
for(i in 1:nrow(y.log)){
  y.chord[i,] = y.log[i,]/rsum[i]
}
```

#### 1) determine an appropriate response model
```{r}
# Determine an appropriate species response model 
y.dca = decorana(y.log)
y.dca
plot(y.dca,type='text')
## Axis length of DCA1 >4 : CCA
## Axis length of DCA1 3-4: CCA & RDA -> Here, I choose to use the RDA.
## Axis length of DCA1 <3 : RDA
```

#### 2) select a set of independent variables
```{r}
# Check multicollinearity in explanatory variables
## found that Temperature and maxT are highly correlated 0.98
round(as.dist(cor(x)),2)    

## Keep the variable with the larger constrained eigenvalue
rda(y.chord~Temperature,data=x) 
rda(y.chord~maxT,data=x)

## Variance inflation factors (VIF)
### Avoid selecting highly correlated independent variables (multicollinearity)
### When R^2 is large, it means that this x variable is not unique -> we dont want.
vif.cca(rda(y.chord~.,data=x))

# Stepwise variable selection: forward # for RDA
y.full = rda(y.chord~.,data=x)
y.red = rda(y.chord~1,data=x)
y.rda = step(y.red,scope=list(lower=~1,upper=formula(y.full))) # step can also be used in regression
y.rda$anova

# Stepwise variable selection: backward
y.rda = step(y.full,scope=list(lower=formula(y.red),upper=formula(y.full)))
y.rda 

# (if intersted)
# Finding the subset of environmental variables so that the Euclidean distances of environmental variable have the max correlation with community dissimilarities
summary(bioenv(y.chord,x,index='euclidean')) 

# Suppose we select these environmental variables
x.new = x[,c(1,10,11)] # I choose three environmental variables with the lowest VIF.
```

#### 3) conduct ordination (RDA or CCA)
```{r}
# Constrained ordination
y.rda = rda(y.chord~.,data=x.new)
summary(y.rda)

# Total inertia = total sum of variance of species columns
sum(apply(y.chord,2,var))

# Compare differences in WA and LC sample scores
plot(y.rda,choices=c(1,2),type='points',display='lc',scaling=1) # either LC or WA or both
points(y.rda,choices=c(1,2),display='wa',pch=19,scaling=1)

# Species and environment correlation for each axis
spenvcor(y.rda)

# IntER- and IntRA-set correlation (structure coefficients)
intersetcor(y.rda)

intrasector = function(object){
  w = weights(object)
  lc = sweep(object$CCA$u,1,sqrt(w),"*")
  cor(qr.X(object$CCA$QR),lc)
}
intrasector(y.rda)

# Triplot
plot(y.rda,choices=c(1,2),display=c('wa','sp','bp'),scaling=2)  #try scaling=1, 2, or 3

plot(y.rda,choices=c(1,2),type='none',scaling=2)   #Use these lines to plot 1 item at a time
points(y.rda,choices=c(1,2),display='wa',pch=19,cex=.5,scaling=2)
text(y.rda,choices=c(1,2),display='sp',col='red',cex=.75,scaling=2)
text(y.rda,choices=c(1,2),display='bp',col='blue')
```

#### 4) interpret the results

#### row 1:10 -> spring
#### row 11:25 -> summer
#### row 26:34 -> winter

###### The results from cluster analysis is similar with RDA. In RDA plot, sites are separated according to species composition and environmental variables. Some sites are closer more than others because they have similar species composition or environmental conditions.
###### "Depth" can well distinguish the data from winter(row26-34) and spring(row1-10). "FishDensity" and "CopepodDensity" can well distinguish the data from summer(row11-25) and others.
```{r}
# Extract the cluster analysis results I've done in previous homework
gp <- c(rep(1,10),rep(2,15),rep(3,9)) 

# Plot the cluster analysis results on the triplot
ordiplot(y.rda)
ordiellipse(y.rda, gp, conf = 0.95, lty = 2, label = T)

ordiplot(y.rda, type = "text")
ordiellipse(y.rda, gp, conf = 0.95, lty = 2)
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 2. Use the “dominant” copepod species data (from HW1) and environmental data (from HW3). Apply partial RDA (or CCA) to quantify how variance of species community is attributed to biological and physical habitat effects. In enviANDdensity.xls, columns B-J are physical variables, and columns K and L are biological variables.

```{r}
# Variance partitioning
y = data # dominant species
x = enviANDdensity
x.physical = x[,c(1:3)] 
x.biological = x[,c(10:11)]

y.log = log(y+1) # log-transform y
y.chord = data.frame(matrix(0,nrow(y.log),ncol(y.log))) #row normalization # chord distance (species data)
rsum = apply(y.log,1,sum)
for(i in 1:nrow(y.log)){
	y.chord[i,] = y.log[i,]/rsum[i]
}

# RDA with plot variables after partialling out patch, landscape, and space effects (modify z matrix for partialling out other effects)
z = cbind(x.physical, x.biological) 
z.plot = rda(y.chord,x,z) # z environment factor that we want to control ( if no 3rd part -> regular RDA) # this is partial RDA

summary(z.plot)
plot(z.plot,choices=c(1,2),display=c('wa','sp','bp'),scaling=2) 


# How variance of species community is attributed to biological and physical habitat effects.
rda <- rda(y.chord ~ ., data = x)
x.physical_rda <- rda(y.chord, x.physical, x.biological)
x.biological_rda <- rda(y.chord, x.biological, x.physical)
x.physical_r2 = x.physical_rda$CCA$tot.chi / x.physical_rda$tot.chi
x.biological_r2 = x.biological_rda$CCA$tot.chi / x.biological_rda$tot.chi
r2 = rda$CCA$tot.chi / rda$tot.chi - x.physical_r2 - x.biological_r2

vp = varpart(y.chord, x.physical, x.biological)
plot(vp, bg = c("red", "blue"), Xnames = c("Physical", "Biological"), alpha = 80)
```





