---
title: "Homework 3"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### Set working directory
```{r}
setwd("~/R")
```

#### Read enviANDdensity.csv into R
```{r}
enviANDdensity <- read.csv(file="~/R/enviANDdensity.csv", header=TRUE)
```
Because we only need the data in "FishDensity..ind..1000m3." and "CopepodDensity.ind..m3.", I take these two columns out and set as "FishDensity" and "CopepodDensity" separately.
```{r}
FishDensity <- enviANDdensity$FishDensity..ind..1000m3.
CopepodDensity <- enviANDdensity$CopepodDensity.ind..m3.
```

√[ Σ(Xi-X̄)2 / (n-1)] 



## 1. Compute the mean and SE(mean) for the fish and copepod densities (assuming all data points are independent) respectively, using both normal theory and non-parametric bootstrap. Plot the histogram of bootstrapped means with bootstrap 1000 times. (Normal theory simply means using the standard formula to estimate SE.)
#### mean and SE(mean)
First, I have to build a SE function in order to process the following steps more easily. 
SE = √[ Σ[(Xi-X̄)＾2] / (n-1)] 
```{r}
StandardError <- function(x){
  square.sum <- 0
  for (i in x){
    square <- (x[i]-mean(x))^2
    square.sum <- square.sum + square
  }
  ( square.sum / (length(x)-1) ) ^ 0.5
}


square.sum <- 0
square <- c()
for (i in FishDensity){
    square[i] <- (FishDensity[i]-mean(FishDensity))^2
}

square
square.sum <- sum(square)

  z <- ( square.sum / (length(FishDensity)-1) ) ^ 0.5
z


square 54438.72

(FishDensity[34]-mean(FishDensity))^2

z <- StandardError(FishDensity)
```

StandarError(FishDensity)



```{r}
FishDensity.mean <- mean(FishDensity)
FishDensity.SE <- var(FishDensity)^0.5/length(FishDensity)^0.5
CopepodDensity.mean <- mean(CopepodDensity)
CopepodDensity.SE <- var(CopepodDensity)^0.5/length(CopepodDensity)^0.5
```
#### non-parametric bootstrap

```{r}
bootstrap.means <- c()
for (i in 1:1000) {
bootstrap.i <- sample(FishDensity, size=length(FishDensity), replace=TRUE)
bootstrap.means[i] <- mean(bootstrap.i)
}
bootstrap.means
```

```{r}
bootstrap.means.mean <- mean(bootstrap.means)
bootstrap.means.mean
```

#### plot the histogram
I plot the histogram of bootstrapped means with bootstrap 1000 times. The black line is the mean of bootstrap means. The blue line is the mean of original sampling. We can see that they are really close to each other.
```{r}
hist(bootstrap.means)
abline(v = FishDensity.mean, col = "black", lwd = 2)
abline(v = bootstrap.means.mean, col = "blue", lwd = 2)
```
Compare the mean of bootstrap means with the original sampling mean.
They are almost the same number, only that through doing bootstrap, our mean of bootstrap means is closer to the original population mean.
```{r}
FishDensity.mean
bootstrap.means.mean
```

## 2. Compute the median and bootstrapped SE(median) for the fish and copepod densities. Plot the histogram of bootstrapped medians with bootstrap 1000 times.
```{r}
FishDensity.median <- median(enviANDdensity$FishDensity..ind..1000m3.)
CopepodDensity.median <- median(enviANDdensity$CopepodDensity.ind..m3.)
```
#### non-parametric bootstrap



bootstrap.medians <- c()
for (i in 1:1000) {
random <- ceiling(runif(34, min=0, max=34))
bootstrap.i <- FishDensity[random]
bootstrap.medians[i] <- median(bootstrap.i)
}
bootstrap.medians



```{r}
bootstrap.medians <- c()
for (i in 1:1000) {
bootstrap.i <- sample(enviANDdensity$FishDensity..ind..1000m3., size=length(enviANDdensity$CopepodDensity.ind..m3.), replace=TRUE)
bootstrap.medians[i] <- median(bootstrap.i)
}
bootstrap.medians
```



```{r}
bootstrap.medians <- c()
for (i in 1:1000) {
bootstrap.i <- sample(enviANDdensity$FishDensity..ind..1000m3., size=length(enviANDdensity$CopepodDensity.ind..m3.), replace=TRUE)
bootstrap.medians[i] <- median(bootstrap.i)
}
bootstrap.medians
```

```{r}
bootstrap.medians.mean <- mean(bootstrap.medians)
bootstrap.medians.mean
```

#### plot the histogram
I plot the histogram of bootstrapped medians with bootstrap 1000 times. The black line is the mean of bootstrap medians. The blue line is the median of original sample. We can see that they are really close to each other. We can hardly see the black line because they are almost ar the same place.
```{r}
hist(bootstrap.medians)
abline(v = FishDensity.median, col = "black", lwd = 2)
abline(v = bootstrap.medians.mean, col = "blue", lwd = 2)
```

Compare the mean of bootstrap medians with the original sampling median.
They are almost the same number, only that through doing bootstrap, our mean of bootstrap medians is closer to the original population median.
```{r}
FishDensity.median
bootstrap.medians.mean
```



## 3. Plot fish (dependent) v.s copepod (independent) and the regression line. Compute the regression coefficients for
fish = β0+β1*copepod and bootstrapped SE(β0) and SE(β1). Plot the histogram of bootstrapped β0 and β1 with bootstrap 1000 times.
```{r}
```




repeat {
 <- ceiling(runif(34, min=0, max=34))
if (x == 6){
break
}
}

repeat
 random = ceoling runif(34, min=0, max=34 )
as.data.frame(rbind(xxx, [random]))
apply # calculate mean by row


sd
standard error of the bootstrap mean

fisfh cop 都要畫圖！


unit results

put originall sample into this  
t(x) %*% x t(x) y


use fynction





A <- as.data.frame()
repeat {
 bootstrap <- ceiling(runif(34, min=0, max=34))
 as.data.fram(rbind(A, [bootstrap]))
 if (length(bootstrap) > 10){
 break
 }
}

A <- as.data.frame()

as.data.frame(x, row.names = NULL, optional = FALSE, …)
```{r}
repeat{
 bootstrap <- ceiling(runif(34, min=0, max=34))
 A <- as.data.frame(rbind( ,[bootstrap]))
 if (length(bootstrap) > 10){
 break
 }
}
```



