---
title: "Homework 6"
author: "Yu-Jen Lin"
output:
  pdf_document: default
  html_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------


#### Set working directory
```{r}
setwd("~/R")
```

#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 1. Write a function to do root-finding based on Newton-Ralphson method and solve for
#### sin(x)/x-0.6=0. (Note: write your own codes, set tol=0.000001, try different initial values)

#### Plot the equation sin(x)/x-0.6=0
I first plot the equation (black) and the horizontal line that y=0 (blue). I find that there are 2 roots of this equation.
```{r}
eq = function(x){y=sin(x)/x-0.6}
curve(eq, from=-50, to=50, xlab="x", ylab="y = sin(x)/x-0.6", ylim = c(-1, 1))
abline(h=0, col=4)


eq = function(x){y=sin(x)/x-0.6}
curve(eq, from=-10, to=10, xlab="x", ylab="y = sin(x)/x-0.6", ylim = c(-1, 1))
abline(h=0, col=4)
```

#### Calculate the derivative of a function in r
In order to calculate the derivative of a function, I install the "Deriv" package.
```{r}
install.packages("Deriv")
library(Deriv)

# Use Deriv, I get the derivative of this function "eq".
# eq = function(x){y=sin(x)/x-0.6}
Deriv(eq) # Through this step, now I know that Deriv(eq) = (cos(x) - sin(x)/x)/x.
```

#### Newton-Ralphson method
#### x1 <- x - f(x)/f'(x)
When improvement < tolerance, stop the loop.
  improvement = | x1 - x | / x  
  tolerance = 0.000001
```{r}
# Set the initial values
x1 = -1 # initial value
x <- 0 # in order to begin the calculation
  
# Use for loop to repeat the process until the tol<=0.000001.
for (i in c(1:100000)) {
  if ((abs(x1-x)/x1) >= 0.000001){  
    # When improvement < tolerance, stop the loop. Now improvement >= tolerance, keep repeating the loop.
    x <- x1
    f.x <- sin(x)/x-0.6
    Fun.deriv.x <- function(x){(cos(x) - sin(x)/x)/x}
  
    # x1 <- x - f(x)/f'(x)
    x1 <- x - f.x / Fun.deriv.x(x)
    
  }else{
    print(x)
    print(i)
    ini.1 <- x
    break
  }
}
```

I repeat last step by setting the different initial values = 2,3 as examples to show that they all converge to the same roots.
```{r}
######## REPEAT ##################################################
# Set the initial values
x1 = 2 # initial value
x <- 0 # in order to begin the calculation
  
# Use for loop to repeat the process until the tol<=0.000001.
for (i in c(1:100000)) {
  if ((abs(x1-x)/x1) >= 0.000001){  
    # When improvement < tolerance, stop the loop. Now improvement >= tolerance, keep repeating the loop.
    x <- x1
    f.x <- sin(x)/x-0.6
    Fun.deriv.x <- function(x){(cos(x) - sin(x)/x)/x}
  
    # x1 <- x - f(x)/f'(x)
    x1 <- x - f.x / Fun.deriv.x(x)
    
  }else{
    ini.2 <- x
    break
  }
}
######## REPEAT ##################################################
# Set the initial values
x1 = 3 # initial value
x <- 0 # in order to begin the calculation
  
# Use for loop to repeat the process until the tol<=0.000001.
for (i in c(1:100000)) {
  if ((abs(x1-x)/x1) >= 0.000001){  
    # When improvement < tolerance, stop the loop. Now improvement >= tolerance, keep repeating the loop.
    x <- x1
    f.x <- sin(x)/x-0.6
    Fun.deriv.x <- function(x){(cos(x) - sin(x)/x)/x}
  
    # x1 <- x - f(x)/f'(x)
    x1 <- x - f.x / Fun.deriv.x(x)
    
  }else{
    ini.3 <- x
    break
  }
}
```

#### All converge to the same roots
I conclude the results and find that different initial values = 1,2,3 all converge to the same roots.
```{r}
initial.x <- c(1,2,3)
roots.x <- c(ini.1, ini.2, ini.3)
roots <- data.frame(initial.x, roots.x)
roots
```


#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### 2. Use data from Vidal (1980) and find the Belehradek's equation for C2, C3, C4, C5 by minimizing the least square error, and set b=-2.05. Plot the data and fitted curves. (Hint: use optim in R or fminsearch in Matlab)

#### Read VidalTvsDuration.txt into R
```{r}
VidalTvsDuration <- read.table(file="~/R/VidalTvsDuration.txt", header=TRUE)
```

#### Belehradek s equation for zooplankton development time
#### D = a*(T-α)^b
D = development time of a stage
a = parameter (unknown)
T = temperature
α = parameter (unknown)
b = -2.05 (we set this parameter)
```{r}
temp <- VidalTvsDuration$tempearture
D.C2 <- VidalTvsDuration$C2
D.C3 <- VidalTvsDuration$C3
D.C4 <- VidalTvsDuration$C4
D.C5 <- VidalTvsDuration$C5




######
data.C2 <- data.frame(D.C2, temp)

FUN.SSR.C2 <- function(data, par){
   with(data.C2, sum (( D.C2 - par[1]*(temp-par[2])^(-2.05) ) ^ 2 ))  # sum of (yi - yi.hat)^2
}
# Minimizing the least square error
optimization <- optim(par=c(5,5), fn=FUN.SSR.C2, data=data.C2)
########




# SSR
FUN.SSR.C2 <- function(a,α){
   SSR.C2   <-   sum (( D.C2 - a*(temp-α)^(-2.05) ) ^ 2 )  # sum of (yi - yi.hat)^2
  return(SSR.C2)
}

# Set the initial values
ini.SSR.C2 <- FUN.SSR.C2(5,7)
ini.SSR.C2








# SSR

FUN.SSR.C2 <- function(a,α){
   SSR.C2   <-   sum (( D.C2 - a*(temp-α)^(-2.05) ) ^ 2 )  # sum of (yi - yi.hat)^2
  return(SSR.C2)
}
# Minimizing the least square error
optimization <- optim(a=1, α=2, fn=FUN.SSR.C2)


#######




# SSR

FUN.SSR.C2 <- function(a,α){
   SSR.C2   <-   sum (( D.C2 - a*(temp-α)^(-2.05) ) ^ 2 )  # sum of (yi - yi.hat)^2
  return(SSR.C2)
}
# Minimizing the least square error
optimization <- optim(c(a=1,α=2), fn=FUN.SSR.C2, data=data.C2)





# Minimizing the least square error
optim(par=c(1,2), fn=FUN.SSR.C2, data=GGG)
u <-optim(par=c(1,2), fn=FUN.SSR.C2)
u


# SSR
SSR.C2 <- 0
FUN.SSR.C2 <- function(a,α){
   SSR.C2   <-   sum (( DDD - a*(TTT-α)^(-2.05) ) ^ 2 )  # sum of (yi - yi.hat)^2
  return(SSR.C2)
}



optim( c(5,7) , FUN.SSR.C2)



# 
plot(
  x = rep(temp,4),
  y = c(D.C2, D.C3, D.C4, D.C5), 
  ylab = "development time",
  xlab = "tempearture")




```
```{r}

F1 <- function(x,y){
  x+2+4*y
  
}

optim(par=c(1,2), fn=F1, data=GGG)
```


