---
title: "Homework 1"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### read the data
First, I have to read the data in R. Thus, I use “read.table” function to read the txt file. The "header=T" means I set the first row as header. On the other hand, the "header=F" means that first row is not a header.
```{r}
composition <- read.table("~/R/copepod_composition.txt", header=T)
head(composition)
density <- read.table("~/R/cop_density.txt", header=F)
head(density)
```

## 1. Calculate the copepod density for each species for each cruise-station

#### transpose
First, in order to make these two dataset multipliable ("composition" and "density"), I do the matrix transpose on the "desity".
```{r}
density.t <- t(density)
```

#### copepod density
Second, I calculate the copepod density by mutiply "density.t" and "composition", and then mutiply 0.01 to make "composition" become a porpotion.
```{r}
density.each <- density.t*composition*0.01
head(density.each)
```

## 2. For each cruise-station, calculate the species richness (number of species) and Shannon diversity index

#### species richness (number of species)
First, I create "density.each.zero" to  distinguish items larger than 0 from the others items. If the items are larger than 0, which means that species does exist in that cruise-station. 
```{r}
density.each.zero <- density.each>0
head(density.each.zero)
```

Second, when I use "apply" to do the FUN (function) by each column (MARGIN=2), I can count the items that are larger than 0 by summing up the "density.each.zero" by row ("True" stands for 1). And then correct it by minus 1.
```{r}
species.number <- apply(density.each.zero, MARGIN=2, FUN = sum)
species.number
species.number.correction <- species.number-1
species.number.correction
```

#### Shannon diversity index
First, I sum up all the density in "density.each" in order to know the total amount of all the species. It's like we assume there's only 1 cubic meter of the sumation of all cruise-stations, and to see the propotion of each species in this 1 cubic meter("density.propotion").
```{r}
density.sum <- sum(density.each)
density.propotion <- density.each/density.sum
head(density.propotion)
```
Second, I mutiply the pi (proportion of individuals belonging to the ith species (wikipedia) ) by log pi, to get the Shannon diversity index for each species. After that, I can easily do the calculation of the real Shannon diversity index. In these process, I do the sumation, use "na.rm=TRUE" to let r not to sum the NAs. Also, I multiply Shannon.sum by -1 because it's in the equation of Shannon diversity index.
```{r}
Shannon <- density.propotion*log(density.propotion)
Shannon.sum <- apply(Shannon, MARGIN=2, FUN = sum, na.rm=TRUE)
Shannon.sum.positive <- Shannon.sum*-1
Shannon.sum.positive
```

## 3. Find dominant species (species >=2% of total composition in any cruise-station) and calculate the average density for the spring, summer, and winter cruise for each dominant species.

#### dominant species
First, I sum up all the propotion in every station ("MARGIN=1" = by row) for each species, and add this vector to the table that contains the density of each species for each cruise-station.
```{r}
density.propotion.sum <- apply(density.propotion, MARGIN=1, FUN = sum) # species sum
density.each.propotion.sum.add <- cbind(density.each,density.propotion.sum)
head(density.each.propotion.sum.add)
```
Second, I pick up the rows that species >=2%.
```{r}
dominant <- subset(density.each.propotion.sum.add, density.propotion.sum >= 0.02, select=c(1:35))
```
Third, I do the sumation for each species in each season. Eventually, use "avg" table to show the results.
```{r}
springsum <- apply(dominant, MARGIN=1, FUN = sum, select=c(1:10) )
springavg <- springsum/10
summersum <- apply(dominant, MARGIN=1, FUN = sum, select=c(11:25) )
summeravg <- summersum/15
wintersum <- apply(dominant, MARGIN=1, FUN = sum, select=c(26:34) )
winteravg <- wintersum/9
avg <- cbind(springavg, summeravg, winteravg)
avg
```
