---
title: "Homework 11"
author: "Yu-Jen Lin"
output:
  html_document: default
  pdf_document: default
---
#### b04b01036@ntu.edu.tw
#### --------------------------------------------------------------------------------------------------------------------------------------------------------

#### The von Bertalanffy Growth Func%on describes the body growth in length of most fish, i.e.

#### L(t)=L∞(1−exp(−K(t−t0)))e^ε       ε~N(0,σ2)
#### L∞ = hypothesized (mean) maximum length
#### K  = growth rate
#### t  = age
#### t0 = hypothesized age when body length is 0

#### Please es%mate the three parameters using the method maximum likelihood for the average length-at-age for male and female Pacific Hake, respec%vely, considering a multiplicative error model with log-normal distribu%on.

## loading data
```{r}
library(knitr)
library(ggplot2)

age = c(1, 2, 3.3, 4.3, 5.3, 6.3, 7.3, 8.3, 9.3, 10.3, 11.3, 12.3, 13.3)
length_m = c(15.4, 26.93, 42.23, 44.59, 47.63, 49.67, 50.87, 52.3, 54.77, 56.43, 55.88)
length_f = c(15.4, 28.03, 41.18, 46.2, 48.23, 50.26, 51.82, 54.27, 56.98, 58.93, 59, 60.91, 61.83)

#show data
kable(cbind(age, length_f, c(length_m, rep("", 2))), col.names = c("Age (year)", "Female mean length (cm)", "Male mean length (cm)"))

```

## Explore the relationship between body length and age
```{r}
#prepare data for plotting
dataset = data.frame(c(age, age[-c(12:13)]), c(length_f, length_m), c(rep("Female", 13), rep("Male", 11)))
colnames(dataset) <- c("Age", "Body length", "Sex")

#draw graph
ggplot(dataset, aes(x = Age, y = `Body length`, col = Sex)) +
    geom_point() +
    scale_color_discrete(name = "Sex")
```
It seems that growth rate declines as age increases and the von Bertalanffy growth function is appropriate.

## The von Bertalanffy growth function
Define function VBGF that calculates estimated length using the von Bertalanffy growth function, which would be used in our regression model.
```{r}
VBGF <-function(x, Linf, K, t0){
    y = Linf * (1 - exp(-K * (x - t0)))
    return(y)
}
```


## Log-likelihood function for the regression
Define function lognormal_like which calculates the log-likelihood of
```{r}
lognormal_like = function(x, y, Linf, K, t0, Out){
    
    like = numeric(length(y))
    NLL = numeric(length(y))
    ypred = VBGF(x, Linf, K, t0)
    dev2 =(log(y) - log(ypred)) ^ 2
    sigma = sqrt(mean(dev2))
    
    for (i in 1:length(y)){
        like[i] = (1 / (x[i] * sqrt(2 * pi) * sigma)) * exp(-dev2[i] / (2 * sigma ^ 2))
        NLL[i] = -log(like[i])
    }
    
    #show each iteration if Out is set as TRUE
    if(Out == T) cat("NLL=", sum(NLL), " Linf=", Linf, " K=", K, " t0=", t0, "\n", sep = "")
    
    #calculate various informations about the log-likelihood
        Outs <- NULL
        Outs$Pred <- ypred
        Outs$Length <- y
        Outs$Dev2 <- dev2
        Outs$sigma <- sigma
        Outs$Like <- like
        Outs$NLL <- NLL 
        Outs$Obj <- sum(NLL)
        
    return(Outs)
}
```

## Define objective function: sum of negative log-likelihood
```{r}
obj_function = function(theda, ...){ 
    Outs <- lognormal_like(Linf = theda[1], K = theda[2], t0 = theda[3], ...)
    obj <- Outs$Obj 
    return(obj)
}
```


## Find the MLE of parameters
Using optimization method, we try to find values of the three parameters (L∞,K,t0
L
∞
K
t
0
) that maximizes the likelihood function (=minimize negative log-likelihood function) for our regression model. Female and male data would be fit seperately.
```{r}
model_f = optim(c(100, 0.2, 0), obj_function, x = age, y = length_f, Out = F, hessian = T)
model_m = optim(c(100, 0.2, 0), obj_function, x = age[-c(12:13)], y = length_m, Out = F, hessian = T)
```

## Display model results
```{r}
param_f <- model_f$par
param_m <- model_m$par
sd_f <- sqrt(diag(solve(model_f$hessian)))
sd_m <- sqrt(diag(solve(model_m$hessian)))
model_result_f <- cbind(param_f, sd_f)
model_result_m <- cbind(param_m, sd_m)
rownames(model_result_f) <- c("L inf", "K", "t0")
rownames(model_result_m) <- c("L inf", "K", "t0")

kable(model_result_f, col.names = c("Estimated parameters", "Standard deviation"), caption = "Model for female fish body length", digits = 3)

kable(model_result_m, col.names = c("Estimated parameters", "Standard deviation"), caption = "Model for male fish body length", digits = 3)
```

## Visualize estimations of regression model
```{r}
#prepare data for plotting
dataset = data.frame(c(age, age[-c(12:13)]), c(length_f, length_m), c(VBGF(age, param_f[1], param_f[2], param_f[3]), VBGF(age[-c(12:13)], param_m[1], param_m[2], param_m[3])), c(rep("Female", 13), rep("Male", 11)))
colnames(dataset) <- c("Age", "Actual length", "Esimated length", "Sex")

#draw graph
ggplot(dataset, aes(x = Age)) +
    geom_point(aes(y = `Actual length`, col = Sex)) +
    geom_line(aes(y = `Esimated length`, col = Sex)) +
    scale_color_discrete(name = "Sex") +
    labs(title = "Regression model using von Bertalanffy growth function", y = "Fish length")
```


```{r}
```


```{r}
```


```{r}
```
